Stochastic Variance-Reduced Policy Gradient

### 2018-02-09 svrpg.tex

\section{Conclusion}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
%{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
%The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
%In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
%}
%\todomatout{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
Preliminary results support the effectiveness of SVRPG also with a commonly used baseline for the policy gradient.
Despite that, we believe that it will be possible to derive a baseline specifically designed for SVRPG in order to jointly exploit the RL structure and the SVRG idea.
\todomat{Other future work}

@@

\section{Practical SVRPG versions}\label{app:practicalsvrpg}
\begin{algorithm}[h]
        \begin{algorithmic}
        \STATE \textbf{Input:} A gradient estimate $g_t$ and parameters $\beta_1$, $\beta_2$, $\epsilon$ and $\alpha$.
        \STATE $m_t = \beta_1 m_{t-1} + (1 - \beta_1) g_t$
        \STATE $v_t = \beta_2 v_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
        \STATE $\hat{m}_t = \dfrac{m_t}{1 - \beta^t_1}$
        \STATE $\hat{v}_t = \dfrac{v_t}{1 - \beta^t_2}$
        \STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t$
        \STATE \textbf{Return:} The increment $\Delta(g_t)$  of the parameters.
\end{algorithmic}
 \caption{
 \label{A:adam}
 Adam}
\end{algorithm}
\subsection{Learning Rates}
Let us give a deeper insight on the two different learning rate used by our algorithm. We have already said that we use two different ADAM one related to the snapshots and another one related to the sub-iterations. In this way tha ADAM associated to the snapshots will take into account only the snapshots history through the parameters $m_{t-1}$ and $v_{t-1}$. The same reasoning holds for the sub-iterations. This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
\subsection{Baseline}
For all batch gradient-based algorithms, we use the same time-varying feature encoding for the linear baseline:
\begin{align*}
\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
\end{align*}
where $s$ is the state vector and $\odot$ is the element-wise product.
\section{Experimental Details}\label{app:exp}
Used environments:
\begin{enumerate}
	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body. Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$. The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
	\item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles. 2-dimensional action space: the two momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$. The episodes terminates when the number of time steps T is greater than 500.
	\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles. 6-dimensional action space: the 6 momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$. The episodes terminates when the number of time steps T is greater than 500.
\end{enumerate}

>>>>

\section{Conclusion}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
%{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
%The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
%In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
%}
%\todomatout{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
Preliminary results support the effectiveness of SVRPG also with a commonly used baseline for the policy gradient.
Despite that, we believe that it will be possible to derive a baseline specifically designed for SVRPG in order to jointly exploit the RL structure and the SVRG idea. Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in \cite{papini2017adaptive}.

@@

\section{Practical SVRPG versions}\label{app:practicalsvrpg}
\begin{algorithm}[h]
        \begin{algorithmic}
        \STATE \textbf{Input:} A gradient estimate $g_t$ and parameters $\beta_1$, $\beta_2$, $\epsilon$ and $\alpha$.
        \STATE $m_t = \beta_1 m_{t-1} + (1 - \beta_1) g_t$
        \STATE $v_t = \beta_2 v_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
        \STATE $\hat{m}_t = \dfrac{m_t}{1 - \beta^t_1}$
        \STATE $\hat{v}_t = \dfrac{v_t}{1 - \beta^t_2}$
        \STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t$
        \STATE \textbf{Return:} The increment $\Delta(g_t)$  of the parameters.
\end{algorithmic}
 \caption{
 \label{A:adam}
 Adam}
\end{algorithm}
\subsection{Learning Rates}
Let us give a deeper insight on the two different learning rate used by our algorithm. We have already said that we use two different ADAM one related to the snapshots and another one related to the sub-iterations. In this way tha ADAM associated to the snapshots will take into account only the snapshots history through the parameters $m_{t-1}$ and $v_{t-1}$. The same reasoning holds for the sub-iterations. This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
\subsection{Baseline}
For all batch gradient-based algorithms, we use the same time-varying feature encoding for the linear baseline:
\begin{align*}
\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
\end{align*}
where $s$ is the state vector and $\odot$ is the element-wise product. In the case of our algorithm the model representing the baseline is fitted every snapshot so we have:\\
\begin{align*}
        \nabla J(\pi^T) =\frac{1}{N} \sum_{n=1}^{N}
                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(y_k^n)\right) (\gamma^h \Reward(y_h^n)-b_{sn})
          + \frac{1}{B} \sum_{n=1}^{B}
                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\\ - \frac{1}{B} \sum_{n=1}^{B}
                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\omega(z_{0:h}|\pi^B,\pi^T)
\end{align*}
Where $b_{sn}$ is the prediction given by the baseline at the snapshot, whereas $b_{si}$ is the prediction given by the baseline at a given subiteration. 
\section{Experimental Details}\label{app:exp}
Used environments:
\begin{enumerate}
	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body. Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$. The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
	\item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles. 2-dimensional action space: the two momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$. The episodes terminates when the number of time steps T is greater than 500.
	\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles. 6-dimensional action space: the 6 momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$. The episodes terminates when the number of time steps T is greater than 500.
\end{enumerate}
